Parsing Tweets into Universal Dependencies

### 2017-12-13 paper.tex
minor change

\section{Introduction}
NLP for social media messages is challenging, both because domain
adaptation is required, but also because creating annotated datasets
(e.g., treebanks)
for training and evaluation is hard. 
Pioneering work by \citet{AAAIW113912} 
annotated 7,630 tokens' worth of tweets according to the
phrase-structure conventions of the Penn Treebank
\citep[PTB;][]{Marcus93buildinga}, enabling conversion to Stanford dependencies.
\citet{kong-EtAl:2014:EMNLP2014} further studied the challenges in
annotating tweets and presented a tweet treebank ({\sc Tweebank}),
consisting of 12,149 tokens and largely following conventions
suggested by \citet{schneider-EtAl:2013:LAW7-ID}, fairly close to 
\citet{Yamada03statisticaldependency} dependencies (without labels). 
Both annotation efforts were highly influenced by the PTB, whose guidelines
have good grammatical coverage on newswire. However, when it comes
to informal, unedited, user-generated text, the guidelines may leave
many annotation decisions unspecified.
%Motivation notes:  explain why UD is attractive (and more attractive
%than what previous work used, which was based on YM).  one reason is
%that it might allow sharing of texts annotated in different genres (or
%even languages, cite Waleed).  but universal annotation schemes are
%challenging!  this paper contends with the challenges of
%applying UD to social media texts.

@@

The regex-based tokenizer of \citet{ICWSM101540}---which was 
originally designed for an exploratory search interface called
TweetMotif, not for NLP---preserves most whitespace-delimited tokens, including 
hashtags, at-mentions, emoticons, and unicode glyphs. 
They also treat contractions and acronyms as whole tokens and do not split them.
UD
tokenization,\footnote{\url{http://universaldependencies.org/u/overview/tokenization.html}}
in order to better serve dependency annotation, treats each syntactic word as a token.
They therefore more aggressively split
clitics from contractions (e.g., {\it  gonna} is tokenized as {\it gon} and {\it na}; {\it its}
is tokenized as {\it it} and {\it s} when {\it s} is a copula).
%\yicomment{are words like gonna called contractions, or do they have
%another term?}. \nascomment{yes}
But acronyms are not touched
in the UD tokenization guidelines. Thus, we follow the UD tokenization for contractions
and leave acronyms like {\em idc} as a single token. 
%\nascomment{unless we make a change, I think we need to say
%  here that acronyms like \emph{idc} are left as a single token}
%\yjcomment{mentioned acronymns, we don't have the problem of normalization, cause
% normalization was just annotated for data study.}

@@

Tokens that abbreviate multiple words, such as \emph{idc} (``I don't
care'') are resolved to the POS of the syntactic head of the
expression, following UD conventions (in this example, the head \emph{care} is
a verb, so \emph{idc} is tagged as a verb).
% The informal syntactic tokens can generally be handled by their distribution.
% However, one of the exception is the {\tt phrasal abbreviations} in which one token carries multiple syntactic functions.
% In dealing with such phrases, we propose a routine that first recover their original forms, then, use the POS of their syntactic head word as the POS of the whole phrase.
% Therefore, {\it idc} (I don't care) is a verb because {\it care} is its syntactic head.
% Such routine works for most of the phrasal abbreviations.
When the token is not phrasal, we use the POS of the left-most
sub-phrase.  For example, \emph{mfw} (``my face when'') is tagged as a
noun (for \emph{face}).  % \footnote{Such failure can be resulted by the mis-match of phrase boundaries between the original form and the definition of `phrase' in UD. For example, `mfw' (my face when) doesn't make a phrase because `when' belongs to the subordination that modifies `my face'.}

@@

\begin{figure*}[t]
	\centering
	\small
	\begin{dependency}[edge slant=2, text only label, label style=above]
		\begin{deptext}
			@username \& its \& \#awesome \& u \& gonna \& $\heart$ \& it \& Chk \& out \& our \& cooool \& project \& on \& http://project\_link \& + \& RT \& pls\\
			\tiny PROPN \& \tiny PRON \& \tiny ADJ \& \tiny PRON \& \tiny VERB \& \tiny VERB \& \tiny PRON \& \tiny VERB \& \tiny ADP \& \tiny PRON \& \tiny ADJ \& \tiny NOUN \& \tiny ADP \& \tiny X \& \tiny SYM \& \tiny VERB\& \tiny INTJ\\
		\end{deptext}
		\deproot{3}{root}
		\depedge[edge unit distance=1em]{3}{1}{vocative}
		\depedge[edge unit distance=1em]{3}{2}{nsubj}
		\depedge[edge unit distance=1em]{3}{5}{parataxis}
		\depedge[edge unit distance=1em]{5}{4}{nsubj}
		\depedge[edge unit distance=1em]{5}{6}{xcomp}
		\depedge[edge unit distance=1em]{6}{7}{obj}
		\depedge[edge unit distance=0.7em]{3}{8}{parataxis}
		\depedge[edge unit distance=1em]{8}{9}{compound:prt}
		\depedge[edge unit distance=0.7em]{8}{12}{obj}
		\depedge[edge unit distance=1em]{12}{10}{nmod:poss}
		\depedge[edge unit distance=1em]{12}{11}{amod}
		\depedge[edge unit distance=0.6em]{8}{14}{obl}
		\depedge[edge unit distance=1em]{14}{13}{case}
		\depedge[edge unit distance=0.55em]{8}{16}{conj}
		\depedge[edge unit distance=1em]{16}{15}{cc}
		\depedge[edge unit distance=0.37em]{3}{17}{discourse}
	\end{dependency}
	\caption{An example tweet with informal but syntactic tokens.
		This is a contrived example inspired by several tweets.}\label{fig:informal-toks}
\end{figure*}

@@

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If one syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. Retweet construction is a special case and we will discuss our choice in the following paragraph;
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one;
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English dataset.
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

Part-of-speech tagging for tweets has been extensively studied \cite{ritter-EtAl:2011:EMNLP,gimpel-EtAl:2011:ACL-HLT2011,derczynski-EtAl:2013:RANLP-2013,owoputi-EtAl:2013:NAACL-HLT,gui-EtAl:2017:EMNLP20172}.
On the annotation scheme designed in \S\ref{sec:pos-anno}, based on UD and adapted for
Twitter, we compared several existing systems which include the Stanford CoreNLP tagger, 
\citet{owoputi-EtAl:2013:NAACL-HLT}'s word cluster enhanced tagger (both in greedy and CRF), and 
also \citet{ma-hovy:2016:P16-1}'s neural network POS tagger 
which has the state-of-the-art performance  on Penn Treebank data.
 %and also two bi-LSTM systems similar
%to \citet{DBLP:journals/corr/HuangXY15} and \citet{lample-EtAl:2016:N16-1}, one using
%word vectors and the other using word and character vectors (see Figure \ref{fig:pos-model}).
All systems were re-trained on the combination of the UD\_English and 
\textsc{Tweebank v2} training sets. We use Twitter-specific glove embeddings released by
\citet{pennington-socher-manning:2014:EMNLP2014} in all neural taggers
and parsers.\footnote{\url{http://nlp.stanford.edu/data/glove.twitter.27B.zip}}

@@

\paragraph{Pipeline evaluation.} Finally, we report  our full
pipeline's performance in  Table \ref{tbl:pipline}. We also compare
our model with a pipeline of the state-of-the-art systems (labeled  ``SOTA''):
Stanford CoreNLP tokenizer, %\yicomment{why not compare with UD pipe?}\yjcomment{i was thinking stanford tokenizer is more suitable, but i will get the udpipe result},
\citet{owoputi-EtAl:2013:NAACL-HLT}'s tagger, and \citet{DBLP:journals/corr/DozatM16}'s parser.
Our system differs from the SOTA pipeline in the tokenization and parser components.
From Table \ref{tbl:pipline}, our pipeline outperforms the SOTA when
evaluated in pipeline manner.
The results also emphasize the
importance of tokenization:  without gold tokenization
%\nascomment{changed ``segmentation'' to ``tokenization'' here, twice},
UD parsing
performance drops by more than four points.

>>>>

\section{Introduction}
NLP for social media messages is challenging, both\yicomment{not only? `both' is still strange to me} because domain
adaptation is required, but also because creating annotated datasets
(e.g., treebanks)
for training and evaluation is hard. 
Pioneering work by \citet{AAAIW113912} 
annotated 7,630 tokens' worth of tweets according to the
phrase-structure conventions of the Penn Treebank
\citep[PTB;][]{Marcus93buildinga}, enabling conversion to Stanford dependencies.
\citet{kong-EtAl:2014:EMNLP2014} further studied the challenges in
annotating tweets and presented a tweet treebank ({\sc Tweebank}),
consisting of 12,149 tokens and largely following conventions
suggested by \citet{schneider-EtAl:2013:LAW7-ID}, fairly close to 
\citet{Yamada03statisticaldependency} dependencies (without labels). 
Both annotation efforts were highly influenced by the PTB, whose guidelines
have good grammatical coverage on newswire. However, when it comes
to informal, unedited, user-generated text, the guidelines may leave
many annotation decisions unspecified.
%Motivation notes:  explain why UD is attractive (and more attractive
%than what previous work used, which was based on YM).  one reason is
%that it might allow sharing of texts annotated in different genres (or
%even languages, cite Waleed).  but universal annotation schemes are
%challenging!  this paper contends with the challenges of
%applying UD to social media texts.

@@

The regex-based tokenizer of \citet{ICWSM101540}---which was 
originally designed for an exploratory search interface called
TweetMotif, not for NLP---preserves most whitespace-delimited tokens, including 
hashtags, at-mentions, emoticons, and unicode glyphs. 
They also treat contractions and acronyms as whole tokens and do not split them.
UD
tokenization,\footnote{\url{http://universaldependencies.org/u/overview/tokenization.html}}
in order to better serve dependency annotation, treats each syntactic word as a token.
They therefore more aggressively split
clitics from contractions (e.g., {\it  gonna} is tokenized as {\it gon} and {\it na}; {\it its}
is tokenized as {\it it} and {\it s} when {\it s} is a copula).
%\yicomment{are words like gonna called contractions, or do they have
%another term?}. \nascomment{yes}
But acronyms are not touched
in the UD tokenization guidelines. Thus, we follow the UD tokenization for contractions
and leave acronyms like {\em idc} (``I don't care'') as a single token. 
%\nascomment{unless we make a change, I think we need to say
%  here that acronyms like \emph{idc} are left as a single token}
%\yjcomment{mentioned acronymns, we don't have the problem of normalization, cause
% normalization was just annotated for data study.}

@@

Tokens that abbreviate multiple words, such as \emph{idc} are resolved to the POS of the syntactic head of the
expression, following UD conventions (in this example, the head \emph{care} is
a verb, so \emph{idc} is tagged as a verb).
% The informal syntactic tokens can generally be handled by their distribution.
% However, one of the exception is the {\tt phrasal abbreviations} in which one token carries multiple syntactic functions.
% In dealing with such phrases, we propose a routine that first recover their original forms, then, use the POS of their syntactic head word as the POS of the whole phrase.
% Therefore, {\it idc} (I don't care) is a verb because {\it care} is its syntactic head.
% Such routine works for most of the phrasal abbreviations.
When the token is not phrasal, we use the POS of the left-most
sub-phrase.  For example, \emph{mfw} (``my face when'') is tagged as a
noun (for \emph{face}).  % \footnote{Such failure can be resulted by the mis-match of phrase boundaries between the original form and the definition of `phrase' in UD. For example, `mfw' (my face when) doesn't make a phrase because `when' belongs to the subordination that modifies `my face'.}

@@

\begin{figure*}[t]
	\centering
	\small
	\begin{dependency}[edge slant=2, text only label, label style=above]
		\begin{deptext}
			@username \& its \& \#awesome \& u \& gonna \& $\heart$ \& it \& Chk \& out \& our \& cooool \& project \& on \& http://project\_link \& + \& RT \& it\\
			\tiny PROPN \& \tiny PRON \& \tiny ADJ \& \tiny PRON \& \tiny VERB \& \tiny VERB \& \tiny PRON \& \tiny VERB \& \tiny ADP \& \tiny PRON \& \tiny ADJ \& \tiny NOUN \& \tiny ADP \& \tiny X \& \tiny SYM \& \tiny VERB\& \tiny PRON\\
		\end{deptext}
		\deproot{3}{root}
		\depedge[edge unit distance=1em]{3}{1}{vocative}
		\depedge[edge unit distance=1em]{3}{2}{nsubj}
		\depedge[edge unit distance=1em]{3}{5}{parataxis}
		\depedge[edge unit distance=1em]{5}{4}{nsubj}
		\depedge[edge unit distance=1em]{5}{6}{xcomp}
		\depedge[edge unit distance=1em]{6}{7}{obj}
		\depedge[edge unit distance=0.7em]{3}{8}{parataxis}
		\depedge[edge unit distance=1em]{8}{9}{compound:prt}
		\depedge[edge unit distance=0.7em]{8}{12}{obj}
		\depedge[edge unit distance=1em]{12}{10}{nmod:poss}
		\depedge[edge unit distance=1em]{12}{11}{amod}
		\depedge[edge unit distance=0.6em]{8}{14}{obl}
		\depedge[edge unit distance=1em]{14}{13}{case}
		\depedge[edge unit distance=0.55em]{8}{16}{conj}
		\depedge[edge unit distance=1em]{16}{15}{cc}
		\depedge[edge unit distance=0.37em]{16}{17}{obj}
	\end{dependency}
	\caption{An example tweet with informal but syntactic tokens.
		This is a contrived example inspired by several tweets.}\label{fig:informal-toks}
\end{figure*}

@@

\begin{itemize}
%\item For the at-mentions in retweet constructions, we attach them to their retweet markers. We will discuss our choice in the following paragraph;
\item If a non-syntactic token is within a sentence that has a clear predicate, it will be attached to this predicate. Retweet construction is a special case and we will discuss our choice in the following paragraph;
\item If the whole sentence is made of a sequence of non-syntactic tokens, we attach all these tokens to the first one;
%\item In parataxis case, when it is difficult to decide which sentence or phrase x should belong to, we always attach x to the previous predicate of the paratactic clause or phrase;
\item Non-syntactic tokens are mostly labeled as {\it discourse}, but
  URLs are always labeled as {\it list}, following the UD\_English dataset.
%\\ \yicomment{For retweet at-mention case, we attach the retweet at-mentions to the RT sign, so that the retweet action and the people whose tweet is being retweeted could be connected.\footnote{We can also follow the first convention to attach the retweet at-mention to its predicate, but we think it is better for downstream tasks.}}
\end{itemize}

@@

Part-of-speech tagging for tweets has been extensively studied \cite{ritter-EtAl:2011:EMNLP,gimpel-EtAl:2011:ACL-HLT2011,derczynski-EtAl:2013:RANLP-2013,owoputi-EtAl:2013:NAACL-HLT,gui-EtAl:2017:EMNLP20172}. We therefore opt for the existing POS taggers for tweets instead of developing our own's.
On the annotation scheme designed in \S\ref{sec:pos-anno}, based on UD and adapted for
Twitter, we compared several existing systems which include the Stanford CoreNLP tagger, 
\citet{owoputi-EtAl:2013:NAACL-HLT}'s word cluster enhanced tagger (both in greedy and CRF), and 
also \citet{ma-hovy:2016:P16-1}'s neural network POS tagger 
which has the state-of-the-art performance  on Penn Treebank data.
 %and also two bi-LSTM systems similar
%to \citet{DBLP:journals/corr/HuangXY15} and \citet{lample-EtAl:2016:N16-1}, one using
%word vectors and the other using word and character vectors (see Figure \ref{fig:pos-model}).
All systems were re-trained on the combination of the UD\_English and 
\textsc{Tweebank v2} training sets. We use Twitter-specific glove embeddings released by
\citet{pennington-socher-manning:2014:EMNLP2014} in all neural taggers
and parsers.\footnote{\url{http://nlp.stanford.edu/data/glove.twitter.27B.zip}}

@@

\paragraph{Pipeline evaluation.} Finally, we report  our full
pipeline's performance in  Table \ref{tbl:pipline}. We also compare
our model with a pipeline of the state-of-the-art systems (labeled  ``SOTA''):
Stanford CoreNLP tokenizer\footnote{We choose Stanford CoreNLP tokenizer in the spirit of comparing rule-based and statistical model-based method.}, %\yicomment{why not compare with UD pipe?}\yjcomment{i was thinking stanford tokenizer is more suitable, but i will get the udpipe result},
\citet{owoputi-EtAl:2013:NAACL-HLT}'s tagger, and \citet{DBLP:journals/corr/DozatM16}'s parser.
Our system differs from the SOTA pipeline in the tokenization and parser components.
From Table \ref{tbl:pipline}, our pipeline outperforms the SOTA when
evaluated in pipeline manner.
The results also emphasize the
importance of tokenization:  without gold tokenization
%\nascomment{changed ``segmentation'' to ``tokenization'' here, twice},
UD parsing
performance drops by more than four points.
